A Monte Carlo Synthetic Acceleration Method for the $SP_N$ Equations

### 2013-03-05 doc/spn/spn.tex
updated spectral analysis plan
finished draft of spectral analysis plan

\subsection{Organization of the Spectral Analysis}
\label{subsec:spectral_organization}
For the spectral analysis we will perform a numerical parameter study
for the multigroup $SP_N$ equations. Based on the above information
for iterative methods, several key quantities should be computed for
the $SP_N$ system:
\begin{itemize}
\item Eigenvalue spectrum for the linear operator $\mathbb{L}$
\item Largest eigenvalue for the linear operator $\mathbb{L}$
\item Smallest eigenvalue for the linear operator $\mathbb{L}$
\item Condition number for the linear operator $\mathbb{L}$
\item For each stationary method:
  \begin{itemize}
  \item Eigenvalue spectrum for the iteration matrix $\mathbf{H}$
  \item Largest eigenvalue for the iteration matrix $\mathbf{H}$
  \item Smallest eigenvalue for the iteration matrix $\mathbf{H}$
  \item Condition number for the iteration matrix $\mathbf{H}$
  \end{itemize}
\item For GMRES:
  \begin{itemize}
  \item Chebyshev polynomial norm using eigenvalues of $\mathbb{L}$
    for a fixed number of iterations determined by stationary method
    convergence
  \end{itemize}
\end{itemize}
In addition, the following parameters may be varied in the $SP_N$
system in Denovo:
\begin{itemize}
  \item $SP_N$ order
  \item Number of energy groups $N_g$
  \item Total cross sections $\sigma^g$
  \item Scattering cross sections $\sigma_{sn}^{gg'}$
  \item Boundary conditions (Marshak or reflecting)
  \item Finite volume mesh size ($\Delta$ in all directions)
  \item Number of mesh elements
  \item Source of neutrons (Uniform or Point)
\end{itemize}
We will restrict the problem description in Denovo to one similar to
Brantley and Larsen's $SP_N$ work where the spectral radius was
computed for convergence analysis a several proposed solution
methods. From this definition we choose a single material problem with
a uniform source of unity. For the boundary conditions, we choose
reflecting conditions on all six sides of the cubic domain for an
effectively homogenous infinite medium problem which we will mesh with
a $5\times 5\times 5$ grid with $\Delta=0.1$. We are then left to vary
the angular and energy compoenents of the system in our parameter
studies: the $SP_N$ order, the number of energy groups, and their
cross sections. In Brantley and Larsen's work, a single energy group
was used and therefore we may not use their cross sections in our
computations.

>>>>

\subsection{Organization of the Spectral Analysis}
\label{subsec:spectral_organization}
For the spectral analysis we will perform a numerical parameter study
for the multigroup $SP_N$ equations. Based on the above information
for iterative methods, several key quantities should be computed for
the $SP_N$ system:
\begin{itemize}
\item Eigenvalue spectrum for the linear operator $\mathbb{L}$
\item Largest eigenvalue for the linear operator $\mathbb{L}$
\item Smallest eigenvalue for the linear operator $\mathbb{L}$
\item Condition number for the linear operator $\mathbb{L}$
\item For each stationary method:
  \begin{itemize}
  \item Eigenvalue spectrum for the iteration matrix $\mathbf{H}$
  \item Largest eigenvalue for the iteration matrix $\mathbf{H}$
  \item Smallest eigenvalue for the iteration matrix $\mathbf{H}$
  \item Condition number for the iteration matrix $\mathbf{H}$
  \end{itemize}
\item For GMRES:
  \begin{itemize}
  \item Chebyshev polynomial norm using eigenvalues of $\mathbb{L}$
    for a fixed number of iterations determined by stationary method
    convergence
  \end{itemize}
\end{itemize}
In addition, the following parameters may be varied in the $SP_N$
system in Denovo:
\begin{itemize}
  \item $SP_N$ (angular flux expansion) order
  \item $P_N$ (scattering cross section expansion) order
  \item Number of energy groups $N_g$
  \item Total cross sections $\sigma^g$
  \item Scattering cross section matrix $\sigma_{sn}^{gg'}$
  \item Boundary conditions (Marshak or reflecting)
  \item Finite volume mesh size ($\Delta$ in all directions)
  \item Number of mesh elements
  \item Source of neutrons (Uniform or Point)
\end{itemize}
We will restrict the problem description in Denovo to one similar to
Brantley and Larsen's $SP_N$ work where the spectral radius was
computed for convergence analysis of several proposed solution
methods. From this definition we choose a single material problem with
a uniform source of unity in all energy groups. For the boundary
conditions, we choose reflecting conditions on all six sides of the
cubic domain for an effectively homogenous infinite medium problem
which we will mesh with a $5\times 5\times 5$ grid with
$\Delta=0.1$. We are then left to vary the angular and energy
components of the system in our parameter studies: the $SP_N$ order,
the number of energy groups, and their cross sections of order
$P_N$. In Brantley and Larsen's work, a single energy group was used
and therefore we may not use their cross sections in our multigroup
computations. For our work, we will always choose a total cross
section of unity for all groups and all flux moments. For the
scattering matrix, two classes will be chosen: one with upscatter and
one without. For the downscatter only case, all groups may scatter
within the group and to all groups of lower energy (yielding a lower
triangular scattering matrix) with a cross section of 0.5. For the
upscatter case, all groups may scatter within the group and to all
groups of lower energy given a cross section of 0.5 and for upscatter,
all groups may scatter up to two groups higher in energy with a cross
section of 0.25. All cross section matrices will be equivalent for
each scattering moment. For these cross sections, the $P_N$ order will
be varied with $N=0,1,3$. Both 1 and 10 energy groups will be used
with the 10 energy group case using the scattering matrices specified
above. For the angular flux, the $SP_N$ order will be varied with
$N=1,3,5,7$. This gives a total of 36 linear system variations for the
spectral analysis.

\subsection{Spectral Analysis Results}
\label{subsec:analysis_results}
